---
title: "writing_functions"
output: github_document
date: "2025-10-23"
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
library(tidyverse)
library(rvest)

set.seed(1)

```

## Start Small
The best way to build up a function is to start with code you’ve written outside a function. 
Simple example: the chunk below takes a sample from a normal distribution and then computes the vector of Z scores for the sample.

```{r}
x_vec = rnorm(20, mean = 10, sd = 3.5)

#computing z scores
(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function to compute z scores

```{r}
#x is a list of numbers
z_scores = function(x) {
 
   z = (x - mean(x)) / sd(x) #compute for x and then call it z
  
   z #often the last line of the function is the output 
}

#z is not being created and stored in the environment. it only exists within this function and when this function is running. 
```


Let's try our function ...

```{r}
z_scores( x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)

```


Let's break our function

```{r, error = TRUE}
z_scores(3)
#breaks because you need at least a couple values to calculate SD

z_scores("my name is jeff")
#breaks becuase you need to use a number to calculate a mean 
```


Update function to let people know when it is breaking...
```{r}
z_scores = function(x) {
 
  if (!is.numeric(x)) {
     stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
     stop("only compute scores where x > 5")
  }
  
  z = (x - mean(x)) / sd(x) #compute for x and then call it z
  
  z #often the last line of the function is the output 
}
```


## let's compute some stuff
What if we want to return multiple values (mean and sd of a numeric vector)

```{r}

mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  else if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)

  c(mean_x, sd_x)
}

mean_and_sd(x_vec)
```

A better output format would be a df

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  else if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)

  tibble(
  mean = mean_x, 
       sd = sd_x
  )
  
}

mean_and_sd(x_vec)
```

Alternative Format

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  else if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)

  list(mean = mean_x, 
       sd = sd_x
  )
  
}

mean_and_sd(x_vec)
```



# Make up data....

Let's *simulate* some data

```{r}
sim_df = 
  tibble(
    x = rnorm(n = 300, mean = 5, sd = 2)
  )

sim_df |> 
  summarize(
    mu_hat = mean(x),
    sigma_hat = sd(x)
    
    )
```


What if i want to createa function that creates this dataframe rather than running the code chunks indvidually multiple times 


Write a function to do simulations
(Originally we wrote this as code chunk. Now, we moved code to a source file)
```{r}
#now i import from source folder:
source("source/sim_mean_sd.R")

#source means execute entire file in backgorund

```


```{r}
sim_mean_sd(n_subj = 30)
```



Let's run this function

```{r}
sim_mean_sd()
sim_mean_sd(n_subj = 3800)
sim_mean_sd(mu = 45)

#Let's say you dont define default value for n_subj
sim_mean_sd = function(n_subj, mu = 3, sigma = 2) { #these are default values if not defined when you actually run the function
  
              sim_df = 
              tibble(
                x = rnorm(n = n_subj, mean = mu, sd = sigma)
              )
          
            sim_df |> 
              summarize(
                mu_hat = mean(x),
                sigma_hat = sd(x)
                )
  
}

sim_mean_sd(50) #it will use 50 for the first non defined value 
sim_mean_sd(mu = 45, 50) #it will use 50 for the first non defined value 
```


Import LoTR data

```{r}
#Warning lights should be going off!
fellowship_ring = 
readxl::read_excel("data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two_towers")

return_king = readxl::read_excel("data/LotR_Words.xlsx", range = "J3:L6") |>
  mutate(movie = "return_king")

lotr_df = 
  bind_rows(fellowship_ring, two_towers, return_king) |>
  janitor::clean_names() |>
  pivot_longer(
    female:male,
    names_to = "sex",
    values_to = "words") |> 
  mutate(race = str_to_lower(race)) |> 
  select(movie, everything()) 
```


Try to import the datasets using a function!

```{r}
#my version
import_lotr = function(path, cells, m) {
  
  df = 
    readxl::read_excel(path, range = cells) |>
    mutate(movie = m) 
  
  df
}

```

r = cells = "B3:D6"
x = pathway ="data/LotR_Words.xlsx"
m = title = "fellowship_ring"


```{r}
#jeff's version
import_lotr = function(cells, movie_title) {
  
  df = 
    readxl::read_excel("data/LotR_Words.xlsx", range = cells) |>
    mutate(movie = movie_title) 
  
  df
}

lotr_df = 
  bind_rows(
    fellowship = import_lotr(cells = "B3:D6", movie_title = "fellowship_ring"),
    two_towers = import_lotr(cells = "F3:H6", movie_title = "two_towers"),
    return_king = import_lotr(cells = "J3:L6", movie_title = "return_king")
  )

```



## Last Example

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>                         #output only the first table
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,                          #pivot everything except state
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


Let's say you want to import multiple tables 

Why dont we write an import function!!

```{r}
nsduh_table = function(html, table_num, table_name) {
  
  table = 
    html |> #replaced with impput value from function 
    html_table() |> 
    nth(table_num) |> #replaced with impput value from function 
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>                             #replaced with impput value from function 
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
  
}


nsduh_table(nsduh_html, table_num = 1, "Table 1")
nsduh_table(nsduh_html, table_num = 2, "Table 2")
```














