---
title: "writing_functions"
output: github_document
date: "2025-10-23"
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
library(tidyverse)
library(rvest)

set.seed(1)

```

## Start Small
The best way to build up a function is to start with code you’ve written outside a function. 
Simple example: the chunk below takes a sample from a normal distribution and then computes the vector of Z scores for the sample.

```{r}
x_vec = rnorm(20, mean = 10, sd = 3.5)

#computing z scores
(x_vec - mean(x_vec)) / sd(x_vec)
```

Write a function to compute z scores

```{r}
#x is a list of numbers
z_scores = function(x) {
 
   z = (x - mean(x)) / sd(x) #compute for x and then call it z
  
   z #often the last line of the function is the output 
}

#z is not being created and stored in the environment. it only exists within this function and when this function is running. 
```


Let's try our function ...

```{r}
z_scores( x = x_vec)

num_vec = rnorm(123, mean = 14, sd = 0.4)

z_scores(x = num_vec)

```


Let's break our function

```{r, error = TRUE}
z_scores(3)
#breaks because you need at least a couple values to calculate SD

z_scores("my name is jeff")
#breaks becuase you need to use a number to calculate a mean 
```


Update function to let people know when it is breaking...
```{r}
z_scores = function(x) {
 
  if (!is.numeric(x)) {
     stop("The input x should be numeric")
  }
  
  if (length(x) < 5) {
     stop("only compute scores where x > 5")
  }
  
  z = (x - mean(x)) / sd(x) #compute for x and then call it z
  
  z #often the last line of the function is the output 
}
```


## let's compute some stuff
What if we want to return multiple values (mean and sd of a numeric vector)

```{r}

mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  else if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)

  c(mean_x, sd_x)
}

mean_and_sd(x_vec)
```

A better output format would be a df

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } 
  
  else if (length(x) < 5) {
    stop("Only compute mean and sd when the input has 5 or more numbers")
  }
  
  mean_x = mean(x, na.rm = TRUE)
  sd_x = sd(x, na.rm = TRUE)

  tibble(
  mean = mean_x, 
       sd = sd_x
  )
  
}

mean_and_sd(x_vec)
```















